---
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "img/README-"
)
```


```{r, echo = FALSE}
library(nlexperiment)
# Set the path to your NetLogo instalation
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
```


## Observations per each simulation run
In this example

* two measures _per simulation run_ are defined (values are reported at the end 
of each simulation run),
* the model will be run repetedly `30` times for every parameter value
* and the model is running with `parallel` option (to save some time)

```{r p3Experiment}
experiment <- nl_experiment(
  model_file = file.path(nl_netlogo_path(), 
                         "models/Sample Models/Earth Science/Fire.nlogo"), 
  while_condition = "any? turtles",
  repetitions = 30,
  run_measures = measures(
    percent_burned = "(burned-trees / initial-trees) * 100",
    progress = "max [pxcor] of patches with [pcolor > 0 and pcolor < 55]"
  ),
  param_values = list(
    density = seq(from = 55, to = 62, by = 1)
  )
)
```

Run the experiment with `parallel` option:

```{r p3Run, cache=TRUE}
result <- nl_run(experiment, parallel = TRUE)
# Join observations with parameter space values:
dat <- nl_get_run_result(result, add_parameters = TRUE)
```

Plot the results - percent burned as a function of density:
```{r p3plot1}
# plot percent burned by density
library(ggplot2)
ggplot(dat, mapping = aes(x = factor(density), y = percent_burned) ) + 
  geom_violin() +
  #geom_jitter(position = position_jitter(width = .1), alpha = 0.3) +
  labs(x = "Forest density", y = "Percent burned")
```

Fire advances from left to right. It is interesting to observe
final fire position (left border = 0 and right = 1) as a function of density.
```{r p3plot2, div=200}
ggplot(dat, mapping = aes(x = factor(density), y = progress/250 + 0.5) ) + 
  geom_jitter(position = position_jitter(width = .1), alpha = 0.3)  +
  theme_minimal() +
  labs(x = "Forest density", y = "Fire progress")
```

