---
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "img/README-"
)
```


```{r, echo = FALSE}
library(nlexperiment)
# Set the path to your NetLogo instalation
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
library(ggplot2)
```


## Reading agent variables
While `run_measures` (see [Observation per run](#observations-per-each-simulation-run)) is good enough for aggregate measures, sometimes we need values for each agent individually.
Example shows usage of `agents_after` parameter to get vertices from Preferential attachment
model.

```{r p6experiment}
experiment <- nl_experiment(
  model_file = file.path(nl_netlogo_path(), 
                         "models/Sample models/Networks/Preferential attachment.nlogo"), 
  iterations = 50,
  agents_after = list(
    vertices = agent_set(
      vars = c("who", "xcor", "ycor"), 
      agents = "turtles"),
    edges = agent_set(
      vars = c(e1 = "[who] of end1", e2 ="[who] of end2"), 
      agents = "links")
  ),
  repetitions = 2,
  random_seed = c(42, 69)
)

```

_Note:_

* _Here random seed is defined as a vector. It is applied to each repetition respectively._
* _Variables in agent set may or may not include variable names._
* _Element `agents_before` is analogous tp `agents_after` - it just gets the data before model runs.)_

Run experiment
```{r p6RunModel, cache=TRUE}
result <- nl_run(experiment)
```

Show graph by using **igraph** package:

```{r p6Sigraph, dpi=200, fig.height=4, cache=FALSE}
library(igraph, quietly = TRUE, warn.conflicts = FALSE)
par(mfrow=c(1,2), mai=c(0,0,0,0))
for(i in 1:experiment$run_options$repetitions) {
  g_edges <- subset(result$agents_after$edges, run_id == i)
  g1 <- graph.data.frame(g_edges, directed = FALSE)
  V(g1)$size <- sqrt(degree(g1))*6
  V(g1)$label <- ifelse(as.numeric(V(g1)$name) < 4, as.numeric(V(g1)$name), NA)
  plot.igraph(g1, margin = 0, 
              vertex.label.font = V(g1)$size * 0.07,
              vertex.label.color = "white",
              vertex.color="darkseagreen4",
              edge.color = "gray",
              vertex.frame.color="#ffffff",
              edge.curved=.1
  )
}
```

_Note:_

* _Data frame `result$agents_after$edges` includes variables from all simulation runs._
* _Use `param_space_id` and/or `run_id` columns to subset individual run or parameter combination_





