---
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "img/README-"
)
```


```{r, echo = FALSE}
library(nlexperiment)
# Set the path to your NetLogo instalation
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
fire_model <- file.path(nl_netlogo_path(), "models/Sample Models/Earth Science/Fire.nlogo")
library(ggplot2)
```


## Parameter space
When parameter space is defined by list of values, it is interpreted as 
all possible combinations of parameters. In this example, model is set
to run with all combinations of `world_size` and `density`:

*Note: `world_size` is a special parameter name. Instead of setting NetLogo variable 
it changes the NetLogo world dimensions.*

```{r p4experiment}
experiment <- nl_experiment(
  model_file = fire_model, 
  while_condition = "any? turtles",
  repetitions = 30,
  run_measures = measures(
    percent_burned = "(burned-trees / initial-trees) * 100",
    progress = "max [pxcor] of patches with [pcolor > 0 and pcolor < 55]"
  ),
  param_values = list(
    world_size = c(100, 250),
    density = seq(from = 56, to = 61)
  )
)

```

We have now `r nrow(experiment$param_space)` rows in the parameter space.
With `r experiment$run_options$repetitions` repetitions this means 
`r nrow(experiment$param_space) * experiment$run_options$repetitions` simulation runs.

Sometimes we do not want to run the model for all combinations of parameter values
(often because it would yield a huge parameter space and hours of waiting for
simulations to end). 

There are several sampling methods (also available as packages in R). See Thiele, Kurth, & Grimm (2014) for [overview](http://jasss.soc.surrey.ac.uk/17/3/11.html). 

In this case use a **data frame with explicit parameter sets** instead of a list in `nl_experiment` or `nl_set_param_space` function. 

Here is an example of parameter space with different density values and
two world sizes where only some of the density values are used for the big sized world:

```{r p4ParamSpace}
# Instead of creating new experiment we can change the original experiment
experiment <- nl_set_param_space( experiment,
  param_values = rbind(
    expand.grid(
      world_size = 100,
      density = seq(from = 56, to = 61)
    ),
    expand.grid(
      world_size = 250,
      density = c(58, 59, 60) # only 3 densities for the big world
    )
  )
)
# Print the parameter space values:
experiment$param_space
```


Run the experiment with the `parallel` attribute:
```{r p4RunModel, cache=TRUE}
result <- nl_run(experiment, parallel = TRUE)
# Join observations with parameter space values:
dat <- nl_get_run_result(result, add_parameters = TRUE)
```

Plot the results - with world size as facets:
```{r p4plot, warning=FALSE}
ggplot(dat, mapping = aes(x = factor(density), y=percent_burned, 
                          color = factor(density %in% c(58, 59, 60)))) + 
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(position = position_jitter(width = .1), alpha = 0.3) +
  facet_grid(. ~ world_size, scales = "free_x", space = "free_x") +
  theme(legend.position="none") +
  scale_color_manual(values = c("gray", "black")) +
  ggtitle("Percent burned as a function of density and world size") +
  xlab("Forest density") +
  ylab("Percent burned")

```


