---
title: "Agents and Patches"
---

```{r ch_opt, child="include/chunk_options.rmd", cache=FALSE}
```


```{r, echo = FALSE}
library(nlexperiment)
# Set the path to your NetLogo instalation
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
```

While `step_measures` and `run_measures` 
(see [Fire experiments](fire.html)) 
is good enough for aggregate measures, sometimes we need to measure 
values for each agent individually.

## Turtles and Links
Example below shows usage of `agents_after` element to read 
the network structure from 
Netlogo Preferential attachment
model (Wilensky, 2005).

<aside>
Wilensky, U. (2005). NetLogo 
[Preferential Attachment model](http://ccl.northwestern.edu/netlogo/models/PreferentialAttachment) . 
Center for Connected Learning and Computer-Based Modeling, Northwestern University, 
Evanston, IL.
</aside>

To get data about agents use `agents_after` argument
with a list of agents variables and type of agent (turtles and links in this example).

Element `agents_before` is analogous tp `agents_after` - it just gets the data before model runs.

```{r agentsModel, title = "Define the experiment"}
experiment <- nl_experiment(
  model_file = 
    file.path(nl_netlogo_path(), 
              "models/Sample models/Networks/Preferential attachment.nlogo"), 
  iterations = 50,
  agents_after = list(
    vertices = agent_set(
      vars = c("who", "xcor", "ycor"), 
      agents = "turtles"),
    edges = agent_set(
      vars = c(e1 = "[who] of end1", e2 ="[who] of end2"), 
      agents = "links")
  ),
  repetitions = 2,
  random_seed = c(42, 69)
)
```
<aside>
Instead of `while_condition` the end of simulation run is here defined 
by `iterations` element (number of simulation steps).

With `repetitions` the number of simulation runs (with same parameters) are defined.
The random seed is defined as a vector. It is applied to each repetition respectively.
</aside>

```{r agentsRun, title = "Run the experiment", cache=TRUE}
result <- nl_run(experiment)
```


The data about individual agents are now stored in `result$agents_after$edges`
and `result$agents_after$vertices`. To join the data with parameter sets use 
`nl_get_result(result, type = "agents_after", sub_type = "edges")`

With **igraph** package one can recreate the graph structure from
agents' data:

```{r agentsGraph, title = "Create graph objects", dpi=200, fig.height=4}
library(igraph, quietly = TRUE, warn.conflicts = FALSE)
glist <- 
  lapply(
    1:experiment$run_options$repetitions,
    function(i){
      g_edges <- subset(result$agents_after$edges, run_id == i)
      g1 <- graph.data.frame(g_edges, directed = FALSE)
      V(g1)$size <- sqrt(degree(g1))*6
      V(g1)$label <- 
        ifelse(as.numeric(V(g1)$name) < 4, as.numeric(V(g1)$name), NA)
      g1
    }
)
```
<aside>
Here we set the labels to the first 4 vertices (measured when created)
and the size of the vertice is set to its degree.
</aside>


```{r agentsGraphPlot, title = "Plot the graphs", dpi=200, fig.height=4}
par(mfrow=c(1,2), mai=c(0,0,0,0))
for(g1 in glist) {
  plot.igraph(g1, margin = 0, 
              vertex.label.font = V(g1)$size * 0.07,
              vertex.label.color = "white",
              vertex.color="darkseagreen4",
              edge.color = "gray",
              vertex.frame.color="#ffffff",
              edge.curved=.1
  )
}
```



## Patches
Getting patches information is analogous to reading agents' data.
Example below shows usage of `patches_after` element to read 
the fur patterns from 
Netlogo Fur model (Wilensky, 2003).

<aside>
Wilensky, U. (2003). NetLogo [Fur model](http://ccl.northwestern.edu/netlogo/models/Fur). 
Center for Connected Learning and Computer-Based Modeling, 
Northwestern University, Evanston, IL.
</aside>

```{r agentsFurModel, title = "Define the model"}
experiment2 <- nl_experiment( 
  model_file = file.path(nl_netlogo_path(), 
                         "models/Sample Models/Biology/Fur.nlogo"), 
  iterations = 20,                                     
  
  param_values = list(
    ratio = c(0.3, 0.35, 0.4, 0.45)
  ),
  patches_after = list(
    patches = patch_set(
      vars = c("pxcor", "pycor", "pcolor"),
      patches = "patches"
    )
  ),
  random_seed = 2
)
```

```{r agentsFurRun, title = "Run the model", cache=TRUE}
result2 <- nl_run(experiment2)
```

```{r agentsFurGet, title = "Get data"}
dat <- nl_get_result(result2, type = "patches_after", sub_type = "patches")
str(dat, vec.len = 2)
```
<aside>
Function `nl_show_patches` uses `ggplot2::geom_raster` to plot patches.
The color palette must be provided by user.
</aside>


```{r agentsFurPlot, title = "Plot results", fig.height=3}
library(ggplot2)
nl_show_patches(result2, x_param = "ratio") +
  ggplot2::scale_fill_manual(values = c("black","white"))
```

## See also

[Ants](ants.html) example demonstrates simple parameter sets 
definition and parameter mapping.

<asideclose>
Get [rmarkdown source](`r knitr::current_input()`) of this page.
</asideclose>
