---
title: "Fire Experiment"
---

```{r ch_opt, child="include/chunk_options.rmd", cache=FALSE}
```


This example uses NetLogo Fire model (Wilensky 1997) to demonstrate
how to define and run experiments,
export view images and
get observations based on measures. 

<aside>
Wilensky, U. (1997). NetLogo Fire model. 
http://ccl.northwestern.edu/netlogo/models/Fire. 
Center for Connected Learning and Computer-Based Modeling, 
Northwestern University, Evanston, IL.
</aside>


```{r Initialize, title = "Setup"}
library(nlexperiment)
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
```
<aside>You have to do this only once per R session.</aside>


## Run model and export view image
To run a NetLogo model simulation we have to create an experiment object
with `nl_experiment` function.
Experiment definition includes a NetLogo model file, 
declares how to run the model,
defines parameter values 
and
what outputs are expected from simulation runs.

For example:

```{r fireModel, title = "Define the experiment"}
experiment <- nl_experiment(
  model_file = "models/Sample Models/Earth Science/Fire.nlogo", 
  param_values = list(density = c(57, 59, 61)),
  while_condition = "any? turtles",
  random_seed = 1,
  export_view = TRUE
)
```
<aside>
The model will run for three parameter values. Each simulation will run as long as there are any turtles.

The `export_view` option enables exporting view after each simulation.
</aside>

```{r fireRun, title = "Run the experiment", cache=TRUE}  
result <- nl_run(experiment)     
```

Find paths to the exported view image files in `result$export` 
or just display them by calling `nl_show_views_grid` function:


```{r fireView, title = "Show exported images", fig.height=3, fig.cap="Images taken at the end of each simulation"}
library(ggplot2)
nl_show_views_grid(result, x_param = "density")
```


## Temporal Measures
Temporal measures quantify temporal behavior of agent based model.
Example below shows how to measure _percent of burned trees_ on each simulation step:

$$ percent_{burned}(t) = \frac{trees_{burned}(t)}{trees_{initial}} \times 100 $$

To define a temporal measure use `step_measures` element.
Measures have to be accurate NetLogo reporters. 


```{r fireStepModel, title = "Define the experiment"}
experiment <- nl_experiment( 
  model_file = "models/Sample Models/Earth Science/Fire.nlogo", 
  while_condition = "any? turtles",
  param_values = list(density = c(57, 59, 61)),
  random_seed = 1,
  step_measures = measures(
    percent_burned = "(burned-trees / initial-trees) * 100"
  )
)
```
<aside>
Note that `burned-trees` and `initial-trees` are global variables in NetLogo 
[Fire model](https://github.com/NetLogo/models/blob/master/Sample Models/Earth Science/Fire.nlogo).
</aside>


```{r fireStepRun, title = "Run the experiment", cache=TRUE}
result <- nl_run(experiment)  
```

To get the observations from the result object
use `nl_get_step_result`:


```{r fireStepResult, title = "Get observations", fig.height=4}
dat <- nl_get_step_result(result)
```
<aside>
Values `run_id` and `step_id` are included in the `result$step` by default.

Parameter values are included only by reference to `parameter_set_id`. The 
function `nl_get_result` joins parameter sets with observation data.
</aside>


```{r fireShowTable}
htmlTable::htmlTable(head(dat))
```
<aside>
Table: first 6 observations. 
</aside>
&nbsp;



```{r fireStepPlot, title = "Plot observations", fig.height=4, fig.cap="Percent of burned trees as a function of time"}
nl_show_step(result, x_param = "density")
```


## Observations per each simulation run
Stochastic models are usually run several times repeteadly for each
parameter set and measured at the end of each run.

In this example we observe percent of burned trees and fire progress
for different `density` parameter. The model will run repetedly `30` times
for every parameter value.


```{r fireRunModel, title = "Define the experiment"}
experiment <- nl_experiment(
  model_file = "models/Sample Models/Earth Science/Fire.nlogo", 
  while_condition = "any? turtles",
  run_measures = measures(
    percent_burned = "(burned-trees / initial-trees) * 100",
    progress = "max [pxcor] of patches with [pcolor > 0 and pcolor < 55]"
  ),
  repetitions = 30,
  param_values = list(
    density = seq(from = 55, to = 62, by = 1)
  )
)
```
<aside>
Use `repetitions` parameter to set the number of simulation
runs for each parameter set

Note that one could define `run_measures` and `step_measures` in the same
experiment.
</aside>

When running experiments with several repetitions we can save some time
with `parallel` option.


```{r fireRunRun, title = "Run the experiment", cache=TRUE}
result <- nl_run(experiment, parallel = TRUE)  
```
<aside>To run experiment on several processor cores, run with `parallel = TRUE` option.</aside>


```{r fireRunResult, title = "Get data"}
dat <- nl_get_run_result(result)
```

One way to show the percent burned as a function of density is to use
violin plot:

```{r fireRunPlot, title = "Plot the results", fig.height=5, fig.width=8, fig.cap="Distribution of final percent of burned trees at different values of forest density."}
library(ggplot2)
ggplot(dat, mapping = aes(x = factor(density), y = percent_burned) ) + 
  geom_violin(fill = "gray90") +
  labs(x = "Forest density", y = "Percent burned") + 
  theme_minimal()
```

Fire advances from left to right. It is interesting to observe
final fire position (left border = 0 and right = 1) as a function of density.


```{r fireRunPlot2, div=200, title = "Plot the results", fig.cap="Distribution of final fire position at different values of forest densities"}
ggplot(dat, mapping = aes(x = factor(density), y = progress/250 + 0.5) ) + 
  geom_boxplot()  +
  #geom_jitter(position = position_jitter(width = .1), alpha = 0.3)  +
  labs(x = "Forest density", y = "Fire progress") +
  theme_minimal()
```


## More
Try other examples:

* [Agents and Patches](agents.html) to see how to observe agents and patches
* [Ants](ants.html) to see how to define simple parameter sets and mapping

<asideclose>
Get [rmarkdown source](`r knitr::current_input()`) of this page.
</asideclose>
