---
title: "**<nl>nl</nl><experiment>experiment</experiment>**"
---

```{r ch_opt, child="include/chunk_options.rmd", cache=FALSE}
```


R package for __exploration__ of  **NetLogo**^1^ agent based models.
<asideclose>
^1^ Wilensky, U. (1999). NetLogo. http://ccl.northwestern.edu/netlogo/. Center for Connected Learning and Computer-Based Modeling, Northwestern University, Evanston, IL.
</asideclose>

It provides functions to
define a complete __experiment__ with parameter sets, measures and run options
in __concise__ structure. 
The cycle of experiment definition, data analysis, visualisations and
parameter fitting can be easily turned 
into __readable__ and __reproducible__ documents.


__RNetLogo__ package^2^ is used as an interface to NetLogo environment.
<asideclose>
^2^ Jan C. Thiele (2014). R Marries NetLogo: Introduction to the RNetLogo Package. Journal
of Statistical Software, 58(2), 1-41. URL http://www.jstatsoft.org/v58/i02/.
</asideclose>

## Installation
```{r eval=FALSE}
install.packages("RNetLogo")
devtools::install_github("bergant/nlexperiment")
```

## Example
Experiment with NetLogo Fire model^3^.

<aside>
^3^ Wilensky, U. (1997). 
[NetLogo Fire model](http://ccl.northwestern.edu/netlogo/models/Fire). 
Center for Connected Learning and Computer-Based Modeling, 
Northwestern University, Evanston, IL.
</aside>

```{r Initialize, title = "Setup"}
library(nlexperiment)
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
```
<aside>Set path to your NetLogo installation.
You have to do this only once per R session.</aside>

```{r fireStepModelIndex, title = "Define the experiment"}
experiment <- nl_experiment(
  model_file = "models/Sample Models/Earth Science/Fire.nlogo", 
  while_condition = "any? turtles",
  param_values = list(density = c(57, 59, 61)),
  random_seed = 1,
  step_measures = measures(
    percent_burned = "(burned-trees / initial-trees) * 100"
  )
)
```
<aside>
Note that `burned-trees` and `initial-trees` are global variables in NetLogo 
[Fire model](https://github.com/NetLogo/models/blob/master/Sample Models/Earth Science/Fire.nlogo).
</aside>

```{r fireStepRunIndex, title = "Run the experiment", cache=TRUE}
result <- nl_run(experiment)
```

```{r fireStepPlotI, title = "Plot observations", fig.height=2.8, fig.width=4, cache=TRUE, fig.cap="Percent of burned trees as a function of time", marginfigure = TRUE}
nl_show_step(result, x = "step_id", y = "percent_burned", 
             x_param = "density")
```


## Learning

Start with examples in the [tutorial](tutorial.html).

For reference use R help by typing `?nlexperiment` or
see package [documentation online](help/index.html).

To learn how to create documents from R code see
[R Markdown — Dynamic Documents for R](http://rmarkdown.rstudio.com/)
and the 
rmarkdown [source files](https://github.com/bergant/nlexperiment/tree/gh-pages) of the 
nlexperiment examples.

## Related
Links to related resources:

nlexperiment requires __RNetLogo__ to connect to NetLogo
https://cran.r-project.org/web/packages/RNetLogo/index.html

<aside>
Jan C. Thiele (2014). R Marries NetLogo: Introduction to the RNetLogo Package. Journal
of Statistical Software, 58(2), 1-41. URL http://www.jstatsoft.org/v58/i02/.
</aside> 

Basic ideas and principles of NetLogo experiment definition is taken from
  the NetLogo's **Behavior Space** 
  http://ccl.northwestern.edu/netlogo/docs/behaviorspace.html 
  and **BehaviorSearch** http://www.behaviorsearch.org/

<aside>
Wilensky, U., & Shargel, B. (2002). 
[BehaviorSpace](http://ccl.northwestern.edu/netlogo/behaviorspace.html). 
Center for Connected Learning and Computer Based Modeling, 
Northwestern University, Evanston, IL. 

Stonedahl, F.: [BehaviorSearch](http://www.behaviorsearch.org/) [computer software]. 
Center for Connected Learning and Computer Based Modeling, 
Northwestern University, Evanston, IL.  (2010)

Thiele, J. C., Kurth, W., & Grimm, V. (2014). Facilitating Parameter Estimation and Sensitivity Analysis of Agent-Based Models: A Cookbook Using NetLogo and R. Journal of Artificial Societies and Social Simulation, 17(3), 11. http://jasss.soc.surrey.ac.uk/17/3/11.html
</aside>

The parallel implementation of `nl_run` function was inspired by
  [RNetLogo vignette](https://cran.rstudio.com/web/packages/RNetLogo/vignettes/parallelProcessing.pdf)

Some examples in the tutorial follow examples and methods from Thiele, 
  Kurth & Grimm (2014) article and its 
  [Supplementary Material](http://sourceforge.net/projects/calibrationsensitivityanalysis/)

This site was created with __knitr__ and __markdown__ packages
  as proposed in http://rmarkdown.rstudio.com/html_document_format.html

Plots were created with __ggplot2__ 
  (http://ggplot2.org) and
  __igraph__ (http://igraph.org/r/)

