
---
title: "**<nl>nl</nl><experiment>experiment</experiment>**"
---


```{r ch_opt, child="include/chunk_options.rmd", cache=FALSE}
```


<experiment>__R package for exploration of  NetLogo agent based models.__</experiment>
<asideclose>
Wilensky, U. (1999). NetLogo. http://ccl.northwestern.edu/netlogo/. Center for Connected Learning and Computer-Based Modeling, Northwestern University, Evanston, IL.
</asideclose>

A tool for NetLogo experiment definition, 
  exploring simulation results and model optimization.
  Makes it easy to turn the cycle of experiment definition, 
  data analysis, visualisations and
  parameter fitting into readable and reproducible documents.

## Installation
```{r eval=FALSE}
devtools::install_github("bergant/nlexperiment")
```


```{r ch_opt, child="tutorial_include.rmd", cache=FALSE}
```


## Package Documentation

Use R help for function reference. Download documentation in [pdf](help/nlexperiment.pdf)
or use package [documentation online](help/index.html).


## Related

__<nl>nl</nl><experiment>experiment</experiment>__ requires [RNetLogo](http://rnetlogo.r-forge.r-project.org/) (Thiele 2014) to connect to NetLogo.
See also [Agent Based Models and RNetLogo](http://blog.revolutionanalytics.com/2014/07/agent-based-models-and-rnetlogo.html)
blog from Joseph Rickert.

<aside>
Jan C. Thiele (2014). R Marries NetLogo: Introduction to the RNetLogo Package. Journal
of Statistical Software, 58(2), 1-41. URL http://www.jstatsoft.org/v58/i02/.
</aside> 

Basic ideas and principles of NetLogo experiments are taken from
  the NetLogo's 
  [Behavior Space](http://ccl.northwestern.edu/netlogo/docs/behaviorspace.html)
  (Wilensky & Shargel 2002)
  and [BehaviorSearch](http://www.behaviorsearch.org/) software
  (Stonedahl 2010). 

<aside>
Wilensky, U., & Shargel, B. (2002). 
[BehaviorSpace](http://ccl.northwestern.edu/netlogo/behaviorspace.html). 
Center for Connected Learning and Computer Based Modeling, 
Northwestern University, Evanston, IL. 

Stonedahl, F.: [BehaviorSearch](http://www.behaviorsearch.org/) [computer software]. 
Center for Connected Learning and Computer Based Modeling, 
Northwestern University, Evanston, IL.  (2010)

Thiele, J. C., Kurth, W., & Grimm, V. (2014). Facilitating Parameter Estimation and Sensitivity Analysis of Agent-Based Models: A Cookbook Using NetLogo and R. Journal of Artificial Societies and Social Simulation, 17(3), 11. http://jasss.soc.surrey.ac.uk/17/3/11.html

Stonedahl, F. J. (2011). Genetic algorithms for the exploration of parameter spaces in agent-based models. PhD Thesis, Northwestern University of Illinois

Alden, K., Read, M., Timmis, J., Andrews, P. S., Veiga-Fernandes, H., & Coles, M. (2013). Spartan: A Comprehensive Tool for Understanding Uncertainty in Simulations of Biological Systems. PLoS Comput Biol, 9(2), e1002916. http://doi.org/10.1371/journal.pcbi.1002916
</aside>


Some examples in the [tutorial](tutorial.html) follow the 
  [RNetLogo Cookbook](http://jasss.soc.surrey.ac.uk/17/3/11.html) from Thiele, 
  Kurth & Grimm (2014) and its 
  [Supplementary Material](http://sourceforge.net/projects/calibrationsensitivityanalysis/).
   
Experiments based on NetLogo flocking model and related measure definitions are 
inspired by (Stonedahl, 2011).

See also [spartan](https://www.york.ac.uk/computational-immunology/software/spartan/)
(Alden et al., 2013), an R package with statistical techniques for robustness and sensitivity analysis which can be used with NetLogo. 


