---
title: "**<nl>nl</nl><experiment>experiment</experiment>**"
---

```{r ch_opt, child="include/chunk_options.rmd", cache=FALSE}
```


<experiment>__R package for exploration of  NetLogo agent based models.__</experiment>
<asideclose>
Wilensky, U. (1999). NetLogo. http://ccl.northwestern.edu/netlogo/. Center for Connected Learning and Computer-Based Modeling, Northwestern University, Evanston, IL.
</asideclose>

A tool for NetLogo experiment definition, 
  exploring simulation results and model optimization.
  Makes it easy to turn the cycle of experiment definition, 
  data analysis, visualisations and
  parameter fitting into readable and reproducible documents.

## Installation
```{r eval=FALSE}
devtools::install_github("bergant/nlexperiment")
```
<!--
## Example
Experiment with NetLogo Fire model^2^.

<aside>
^2^ Wilensky, U. (1997). 
[NetLogo Fire model](http://ccl.northwestern.edu/netlogo/models/Fire). 
Center for Connected Learning and Computer-Based Modeling, 
Northwestern University, Evanston, IL.
</aside>

```{r Initialize, title = "Setup"}
library(nlexperiment)
nl_netlogo_path("c:/Program Files (x86)/NetLogo 5.2.0/") 
```
<aside>Set path to your NetLogo installation.
You have to do this only once per R session.</aside>

```{r fireStepModelIndex, title = "Define the experiment"}
experiment <- nl_experiment(
  model_file = "models/Sample Models/Earth Science/Fire.nlogo", 
  while_condition = "any? turtles",
  param_values = list(density = c(57, 59, 61)),
  random_seed = 1,
  step_measures = measures(
    percent_burned = "(burned-trees / initial-trees) * 100"
  )
)
```
<aside>
Note that `burned-trees` and `initial-trees` are global variables in NetLogo 
[Fire model](https://github.com/NetLogo/models/blob/master/Sample Models/Earth Science/Fire.nlogo).
</aside>

```{r fireStepRunIndex, title = "Run the experiment", cache=TRUE}
result <- nl_run(experiment)
```

```{r fireStepPlotI, title = "Plot observations", fig.height=2.8, fig.width=4, cache=TRUE, fig.cap="Percent of burned trees as a function of time", marginfigure = TRUE}
nl_show_step(result, x = "step_id", y = "percent_burned", 
             x_param = "density")
```

-->

## Getting Started

Start with examples in the __[tutorial](tutorial.html)__.

## Package Documentation

For reference use R help, download documentation in [pdf](help/nlexperiment.pdf)
or use package [documentation online](help/index.html).


## Related

__<nl>nl</nl><experiment>experiment</experiment>__ requires [RNetLogo](http://rnetlogo.r-forge.r-project.org/) (Thiele 2014) to connect to NetLogo.
See also [Agent Based Models and RNetLogo](http://blog.revolutionanalytics.com/2014/07/agent-based-models-and-rnetlogo.html)
blog from Joseph Rickert.

<aside>
Jan C. Thiele (2014). R Marries NetLogo: Introduction to the RNetLogo Package. Journal
of Statistical Software, 58(2), 1-41. URL http://www.jstatsoft.org/v58/i02/.
</aside> 

Basic ideas and principles of NetLogo experiment definition are taken from
  the NetLogo's 
  [Behavior Space](http://ccl.northwestern.edu/netlogo/docs/behaviorspace.html)
  (Wilensky & Shargel 2002)
  and [BehaviorSearch](http://www.behaviorsearch.org/) software
  (Stonedahl 2010). 

<aside>
Wilensky, U., & Shargel, B. (2002). 
[BehaviorSpace](http://ccl.northwestern.edu/netlogo/behaviorspace.html). 
Center for Connected Learning and Computer Based Modeling, 
Northwestern University, Evanston, IL. 

Stonedahl, F.: [BehaviorSearch](http://www.behaviorsearch.org/) [computer software]. 
Center for Connected Learning and Computer Based Modeling, 
Northwestern University, Evanston, IL.  (2010)

Thiele, J. C., Kurth, W., & Grimm, V. (2014). Facilitating Parameter Estimation and Sensitivity Analysis of Agent-Based Models: A Cookbook Using NetLogo and R. Journal of Artificial Societies and Social Simulation, 17(3), 11. http://jasss.soc.surrey.ac.uk/17/3/11.html


</aside>


Some examples in the [tutorial](tutorial.html) follow a RNetLogo Cookbook from Thiele, 
  Kurth & Grimm (2014) and its 
  [Supplementary Material](http://sourceforge.net/projects/calibrationsensitivityanalysis/).




